###Appendix S5
## Testing for the correlation of free-running circannual period with survival and BMR using MCMCglmm and mulTree

Supplementary online material  to Julia Karagicheva, Eldar Rakhimberdiev, Anatoly Saveliev & Theunis Piersma 2017 _Circannual rhythms functionally linking life histories and life cycles in birds._ - Journal of Functional Ecology 000: 000-000.

Here we provide a code for the analysis presented in the paper, assessing the relationships between the circannual periods and species-specific survival rate (Model 1)and BMR (Model 2). 

We ran both MCMCglmm models using the reviewed data on 14 passerine species and 2 shorebird species (Circannual_cycles_birds_data.csv), and then repeated the procedure on the same dataset with shorebirds excluded. 

Then we extended each glmmMCMC model by incorporating phylogenetic variance in it, using the package mulTree. From the output, we calculated phylogenetic signal and an analogue of Pagel lambda.

Install required packages
```{r, eval = F}
install.packages('snow')
if(!require(devtools)) install.packages("devtools")
library(devtools)
install_github("TGuillerme/mulTree", ref = "release")
#you might need to install some packages manually
#install.packages(c("MCMCglmm", "coda", "hdrcde", "snow", "ape", "corpcor", "curl"))
```
Load required packages
```{r, eval = F}
library(snow)
library(mulTree)
```
Because running the four models (Model 1 and Model 2, each with and without shorebirds) takes a long time, you might want to tackle them in separate R windows. Doing so, run thee code below ("PREPARE DATA") before each model.

## PREPARE THE DATA 

Download data from GitHub

load table with circannual cycles and deviations of circannual period from 365 days (main table "Circannual_cycles_birds_data")
```{r, eval = F}
Circannual_cycles_birds_data<-read.csv('https://git.io/v5GIF',stringsAsFactors=F)
```
create unique ID 
```{r, eval = F}
Circannual_cycles_birds_data$uniqueid<-paste(Circannual_cycles_birds_data$species,Circannual_cycles_birds_data$Reference,Circannual_cycles_birds_data$BirdID_char,sep="_")
```
make cycle number character
```{r, eval = F}
Circannual_cycles_birds_data$cyclenr_char<-as.character(Circannual_cycles_birds_data$cyclenr_char)
```
load table "Survival_birds_data" with species-specific survival and calculate median per species
```{r, eval = F}
Survival_birds_data<-read.csv('https://git.io/v5Gtq',stringsAsFactors=F)
surv_sp<-aggregate(Survival_birds_data$survival,by=list(Survival_birds_data$species),median)
```
add survival values to the main table
```{r, eval = F}
Circannual_cycles_birds_data$surv<-NA
for(i in 1:nrow(Circannual_cycles_birds_data)){
Circannual_cycles_birds_data$surv[i]<-unique(surv_sp[surv_sp$Group.1 %in% Circannual_cycles_birds_data$species[i],]$x)
}
```
load table "BMR_birds_data.csv" with species-specific BMR  and calculate median per species
```{r, eval = F}
BMR_birds_data<-read.csv('https://git.io/v5ZSb',stringsAsFactors=F)
bmr_sp_median<-aggregate(BMR_birds_data$BMR_W,by=list(BMR_birds_data$species),median)
```
add BMR values to the main table
```{r, eval = F}
Circannual_cycles_birds_data$bmr<-NA
for(i in 1:nrow(Circannual_cycles_birds_data)){
Circannual_cycles_birds_data$bmr[i]<-unique(bmr_sp_median[bmr_sp_median$Group.1==Circannual_cycles_birds_data$species[i],]$x)
}
```
load table "Body_mass_birds_data" with species-specific body masses
```{r, eval = F}
Body_mass_birds_data<-read.csv('https://git.io/v5Z9m',stringsAsFactors=F)
```
add body mass values to the main table
```{r, eval = F}
Circannual_cycles_birds_data$bodymass<-NA
for(i in 1:nrow(Circannual_cycles_birds_data)){
Circannual_cycles_birds_data$bodymass[i]<-unique(Body_mass_birds_data[Body_mass_birds_data$species==Circannual_cycles_birds_data$species[i],]$bodymass)
}
```
Get residuals of survival on BMR
```{r, eval = F}
Circannual_cycles_birds_data$resbmr<-residuals(lm(log(surv)~log(bmr),data=Circannual_cycles_birds_data))
```
Residuals of deviation of circannual period from 365 days on BMR
```{r, eval = F}
Circannual_cycles_birds_data$res_dev_bmr<-residuals(lm(deviation~log(bmr),data=Circannual_cycles_birds_data))
```

## RUN THE MODELS

## Circannual period and survival rate

### MODEL 1 WITH SHOREBIRDS

### set priors
```{r, eval = F}
p.var<-var(Circannual_cycles_birds_data$res_dev_bmr,na.rm=TRUE)
pri5_1 <- list(R = list(V = matrix(p.var/2), nu = 1), G = list(G1 = list(V = matrix(p.var/2), nu = 1),G2 = list(V = matrix(p.var/2), nu = 1), G3=list(V = matrix(p.var/2), nu = 1), G4=list(V = matrix(p.var/2), nu = 1),G5=list(V = matrix(p.var/2), nu = 1)))#
pri4_1 <- list(R = list(V = matrix(p.var/2), nu = 1), G = list(G1 = list(V = matrix(p.var/2), nu = 1),G2 = list(V = matrix(p.var/2), nu = 1), G3=list(V = matrix(p.var/2), nu = 1), G4=list(V = matrix(p.var/2), nu = 1)))#
```
### Run MCMCglmm without pedigree (phylogenetic variable)
```{r, eval = F}
model1<- MCMCglmm(res_dev_bmr~resbmr:cyclenr_char+cyclenr_char,random= ~uniqueid+Trait+Reference+species,data= Circannual_cycles_birds_data, prior=pri4_1, nitt=130000*10,thin=10*5,burnin=3000*10,pr=TRUE,verbose=F,nodes="ALL")

model2<- MCMCglmm(res_dev_bmr~resbmr+cyclenr_char,random= ~uniqueid+Trait+Reference+species,data= Circannual_cycles_birds_data, prior=pri4_1, nitt=130000*10,thin=10*5,burnin=3000*10,pr=TRUE,verbose=F,nodes="ALL")

model1$DIC
model2$DIC

### The best model
summary(model1)

 # # # # Iterations = 30001:1299951
 # # # # Thinning interval  = 50
 # # # # Sample size  = 25400 

 # # # # DIC: 4000.789 

 # # # # G-structure:  ~uniqueid

         # # # # post.mean l-95% CI u-95% CI eff.samp
# # # # uniqueid     558.7    262.2    880.8    25400

               # # # # ~Trait

      # # # # post.mean l-95% CI u-95% CI eff.samp
# # # # Trait      1099      263     2375    25400

               # # # # ~Reference

          # # # # post.mean l-95% CI u-95% CI eff.samp
# # # # Reference      1379    199.4     3129    26957

               # # # # ~species

        # # # # post.mean l-95% CI u-95% CI eff.samp
# # # # species     670.9    112.9     1566    25400

 # # # # R-structure:  ~units

      # # # # post.mean l-95% CI u-95% CI eff.samp
# # # # units      1282     1059     1520    25400

 # # # # Location effects: res_dev_bmr ~ resbmr:cyclenr_char + cyclenr_char 

                     # # # # post.mean l-95% CI u-95% CI eff.samp   pMCMC   
# # # # (Intercept)             -6.419  -40.786   26.613    25400 0.69850   
# # # # cyclenr_char1           10.411    1.506   19.205    25400 0.02087 * 
# # # # resbmr:cyclenr_char0   196.127   71.361  325.515    25350 0.00315 **
# # # # resbmr:cyclenr_char1   118.994   -1.772  239.422    25547 0.05165 . 
# # # # ---
# # # # Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# # # # > 

# control for autocorrelation

autocorr(model1$VCV)
```

### Run the best model with phylogeny using mulTree

set the directory, where you wish to store the outputs

load phylogenetic trees (containing shorebirds), directly from http://birdtree.org/ or just download the 100 trees used in the presented analysis from https://git.io/v5ZxG :
```{r, eval = F}
download.file("https://raw.githubusercontent.com/eldarrak/FLightR/master/examples/Black-Tailed_Godwit_JAB_example/A2_raw_data.lux", "A2_raw_data.lux")
tr_full<-with_shorebirds_tree_subset
```
# set the model structure and parameters
mulTree_data <- as.mulTree(data = Circannual_cycles_birds_data, tree = tr_full,    taxa = "animal",  rand.term=~uniqueid+Trait+Reference+animal+species)
formula_fintest<-res_dev_bmr~resbmr:cyclenr_char+cyclenr_char
mulTree.parameters<-c(5000000, 5000, 100000)

# run mulTree

surv_w_shore<-mulTree(mulTree.data = mulTree_data, formula = formula_fintest, priors = pri5_1, pr=TRUE, parameters = mulTree.parameters, output = "surv_w_shore", chains=8, parallel='SOCK')
```


### MODEL 1 NO SHOREBIRDS

### remove shorebirds from the data table
```{r, eval = F}
Circannual_cycles_no_shorebirds<-Circannual_cycles_birds_data[!(Circannual_cycles_birds_data$Species %in% c("Calidris canutus", "Calidris tenuirostris")),]
```
### set priors
```{r, eval = F}
p.var<-var(Circannual_cycles_birds_data$res_dev_bmr,na.rm=TRUE)
pri5_1 <- list(R = list(V = matrix(p.var/2), nu = 1), G = list(G1 = list(V = matrix(p.var/2), nu = 1),G2 = list(V = matrix(p.var/2), nu = 1), G3=list(V = matrix(p.var/2), nu = 1), G4=list(V = matrix(p.var/2), nu = 1),G5=list(V = matrix(p.var/2), nu = 1)))#
pri4_1 <- list(R = list(V = matrix(p.var/2), nu = 1), G = list(G1 = list(V = matrix(p.var/2), nu = 1),G2 = list(V = matrix(p.var/2), nu = 1), G3=list(V = matrix(p.var/2), nu = 1), G4=list(V = matrix(p.var/2), nu = 1)))#
```
### Run MCMCglmm without pedigree (phylogenetic variable)
```{r, eval = F}
model1_no<- MCMCglmm(res_dev_bmr~resbmr:cyclenr_char+cyclenr_char,random= ~uniqueid+Trait+Reference+species,data= Circannual_cycles_no_shorebirds, prior=pri4_1, nitt=130000*10,thin=10*5,burnin=3000*10,pr=TRUE,verbose=F,nodes="ALL")

model2_no<- MCMCglmm(res_dev_bmr~resbmr+cyclenr_char,random= ~uniqueid+Trait+Reference+species,data= Circannual_cycles_no_shorebirds, prior=pri4_1, nitt=130000*10,thin=10*5,burnin=3000*10,pr=TRUE,verbose=F,nodes="ALL")

model1_no$DIC
model2_no$DIC

# the best model

summary(model1_no)


# # # Iterations = 30001:1299951
 # # # Thinning interval  = 50
 # # # Sample size  = 25400 

 # # # DIC: 2674.365 

 # # # G-structure:  ~uniqueid

         # # # post.mean l-95% CI u-95% CI eff.samp
# # # uniqueid     762.9    337.5     1196    25400

               # # # ~Trait

      # # # post.mean l-95% CI u-95% CI eff.samp
# # # Trait      1077    211.7     2478    25617

               # # # ~Reference

          # # # post.mean l-95% CI u-95% CI eff.samp
# # # Reference      2144    204.4     5182    25400

               # # # ~species

        # # # post.mean l-95% CI u-95% CI eff.samp
# # # species     922.3    113.5     2382    25400

 # # # R-structure:  ~units

      # # # post.mean l-95% CI u-95% CI eff.samp
# # # units      1344     1028     1681    24765

 # # # Location effects: res_dev_bmr ~ resbmr:cyclenr_char + cyclenr_char 

                     # # # post.mean l-95% CI u-95% CI eff.samp   pMCMC   
# # # (Intercept)            -11.237  -54.014   30.641    24511 0.58433   
# # # cyclenr_char1            8.284   -6.171   22.890    25400 0.26252   
# # # resbmr:cyclenr_char0   222.401   68.197  380.447    25400 0.00457 **
# # # resbmr:cyclenr_char1   110.334  -35.269  263.632    25400 0.13039   
# # # ---
# # # Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

# control for autocorrelation

autocorr(model1_no$VCV)
```
### Run the best model with phylogeny using mulTree

set the directory, where you wish to store the outputs
```{r, eval = F}
setwd("C:/Users/meeT55/Desktop/Julia/analysis_with_bmr/LOG-transformed/after_Martin/surv_no_shorebirds")

#load phylogenetic trees not containing shorebirds

load("C:\\Users\\meeT55\\Desktop\\Julia\\data\\all_birds_tree_subset.RData")
tr_full<-all_birds_tree_subset

# set the model structure and parameters

mulTree_data <- as.mulTree(data = Circannual_cycles_no_shorebirds, tree = tr_full,    taxa = "animal",  rand.term=~uniqueid+Trait+Reference+animal+species)
formula_fintest<-res_dev_bmr~resbmr:cyclenr_char+cyclenr_char
mulTree.parameters<-c(5000000, 5000, 100000)

# run mulTree
surv_n_shore<-mulTree(mulTree.data = mulTree_data, formula = formula_fintest, priors = pri5_1, pr=TRUE, parameters = mulTree.parameters, output = "surv_n_shore", chains=8, parallel='SOCK')
```

## Circannual period and BMR

### MODEL 2 WITH SHOREBIRDS

### residuals of the deviation of circannual cycle from 365 days (the response variable) and log(BMR) on log(survival)
```{r, eval = F}
Circannual_cycles_birds_data$devbysurv<-residuals(lm(deviation~log(surv),data=Circannual_cycles_birds_data))
Circannual_cycles_birds_data$bmrbysurv<-residuals(lm(log(bmr)~log(surv),data=Circannual_cycles_birds_data))
```
### set priors
```{r, eval = F}
p.var<-var(Circannual_cycles_birds_data$devbysurv,na.rm=TRUE)
pri5_1 <- list(R = list(V = matrix(p.var/2), nu = 1), G = list(G1 = list(V = matrix(p.var/2), nu = 1),G2 = list(V = matrix(p.var/2), nu = 1), G3=list(V = matrix(p.var/2), nu = 1), G4=list(V = matrix(p.var/2), nu = 1),G5=list(V = matrix(p.var/2), nu = 1)))#
pri4_1 <- list(R = list(V = matrix(p.var/2), nu = 1), G = list(G1 = list(V = matrix(p.var/2), nu = 1),G2 = list(V = matrix(p.var/2), nu = 1), G3=list(V = matrix(p.var/2), nu = 1), G4=list(V = matrix(p.var/2), nu = 1)))#
```
### Run MCMCglmm without phylogeny
```{r, eval = F}
model1_bmr<- MCMCglmm(devbysurv~bmrbysurv:cyclenr_char+cyclenr_char,random= ~uniqueid+Trait+Reference+species,data= Circannual_cycles_birds_data, prior=pri4_1, nitt=130000*10,thin=10*5,burnin=3000*10,pr=TRUE,verbose=F,nodes="ALL")

model2_bmr<- MCMCglmm(devbysurv~bmrbysurv+cyclenr_char,random= ~uniqueid+Trait+Reference+species,data= Circannual_cycles_birds_data, prior=pri4_1, nitt=130000*10,thin=10*5,burnin=3000*10,pr=TRUE,verbose=F,nodes="ALL")

model1_bmr$DIC
model2_bmr$DIC

#the best model

summary(model1_bmr)
 # # # Iterations = 30001:1299951
 # # # Thinning interval  = 50
 # # # Sample size  = 25400 

 # # # DIC: 4006.412 

 # # # G-structure:  ~uniqueid

         # # # post.mean l-95% CI u-95% CI eff.samp
# # # uniqueid     544.2    249.8    867.7    24843

               # # # ~Trait

      # # # post.mean l-95% CI u-95% CI eff.samp
# # # Trait      1079    251.6     2344    26048

               # # # ~Reference

          # # # post.mean l-95% CI u-95% CI eff.samp
# # # Reference      1317    185.4     3023    24651

               # # # ~species

        # # # post.mean l-95% CI u-95% CI eff.samp
# # # species     675.3    115.3     1571    24807

 # # # R-structure:  ~units

      # # # post.mean l-95% CI u-95% CI eff.samp
# # # units      1304     1079     1543    25400

 # # # Location effects: devbysurv ~ bmrbysurv:cyclenr_char + cyclenr_char 

                        # # # post.mean  l-95% CI  u-95% CI eff.samp  pMCMC   
# # # (Intercept)               -3.7186  -36.5992   29.8792    25400 0.8224   
# # # cyclenr_char1              9.2915    0.2732   18.1324    24378 0.0424 * 
# # # bmrbysurv:cyclenr_char0  -91.8645 -160.9604  -23.3079    25400 0.0100 **
# # # bmrbysurv:cyclenr_char1  -63.3397 -127.7827    1.6500    25400 0.0536 . 
# # # ---
# # # Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1


# control for autocorrelation

autocorr(model1_no$VCV)
```
### Run the best model with phylogeny using mulTree

set the directory, where you wish to store the outputs
```{r, eval = F}
setwd("C:/Users/meeT55/Desktop/Julia/analysis_with_bmr/LOG-transformed/after_Martin/bmr_with_shorebirds")

###load phylogenetic trees containing shorebirds

load("C:\\Users\\meeT55\\Desktop\\Julia\\data\\with_shorebirds_tree_subset.RData")
tr_full<-with_shorebirds_tree_subset

### set the model structure and parameters
mulTree_data <- as.mulTree(data = Circannual_cycles_birds_data, tree = tr_full,    taxa = "animal",  rand.term=~uniqueid+Trait+Reference+animal+species)
formula_fintest<-devbysurv~bmrbysurv:cyclenr_char+cyclenr_char
mulTree.parameters<-c(5000000, 5000, 100000)

# run mulTree
bmr_w_shore<-mulTree(mulTree.data = mulTree_data, formula = formula_fintest, priors = pri5_1, pr=TRUE, parameters = mulTree.parameters, output = "bmr_w_shore", chains=8, parallel='SOCK')
```

### MODEL 2 NO SHOREBIRDS

### residuals of the deviation of circannual cycle from 365 days (the response variable) and log(BMR) on log(survival)
```{r, eval = F}
Circannual_cycles_birds_data$bmrbysurv<-residuals(lm(log(bmr)~log(surv),data=Circannual_cycles_birds_data))
Circannual_cycles_birds_data$devbysurv<-residuals(lm(deviation~log(surv),data=Circannual_cycles_birds_data))
```
### exclude shorebirds from the data table
```{r, eval = F}
Circannual_cycles_no_shorebirds<-Circannual_cycles_birds_data[!(Circannual_cycles_birds_data$Species %in% c("Calidris canutus", "Calidris tenuirostris")),]
```
### set priors
```{r, eval = F}
p.var<-var(Circannual_cycles_birds_data$devbysurv,na.rm=TRUE)
pri5_1 <- list(R = list(V = matrix(p.var/2), nu = 1), G = list(G1 = list(V = matrix(p.var/2), nu = 1),G2 = list(V = matrix(p.var/2), nu = 1), G3=list(V = matrix(p.var/2), nu = 1), G4=list(V = matrix(p.var/2), nu = 1),G5=list(V = matrix(p.var/2), nu = 1)))#
pri4_1 <- list(R = list(V = matrix(p.var/2), nu = 1), G = list(G1 = list(V = matrix(p.var/2), nu = 1),G2 = list(V = matrix(p.var/2), nu = 1), G3=list(V = matrix(p.var/2), nu = 1), G4=list(V = matrix(p.var/2), nu = 1)))#
```
### Run MCMCglmm without phylogeny
```{r, eval = F}
model1_bmr<- MCMCglmm(devbysurv~bmrbysurv:cyclenr_char+cyclenr_char,random= ~uniqueid+Trait+Reference+species,data= Circannual_cycles_no_shorebirds, prior=pri4_1, nitt=130000*10,thin=10*5,burnin=3000*10,pr=TRUE,verbose=F,nodes="ALL")
model2_bmr<- MCMCglmm(devbysurv~bmrbysurv+cyclenr_char,random= ~uniqueid+Trait+Reference+species,data= Circannual_cycles_no_shorebirds, prior=pri4_1, nitt=130000*10,thin=10*5,burnin=3000*10,pr=TRUE,verbose=F,nodes="ALL")

model1_bmr$DIC
model2_bmr$DIC

# the best model

summary(model1_bmr)

 # # # Iterations = 30001:1299951
 # # # Thinning interval  = 50
 # # # Sample size  = 25400 

 # # # DIC: 2673.706 

 # # # G-structure:  ~uniqueid

         # # # post.mean l-95% CI u-95% CI eff.samp
# # # uniqueid     761.4    333.6     1194    25375

               # # # ~Trait

      # # # post.mean l-95% CI u-95% CI eff.samp
# # # Trait      1069    196.9     2444    25400

               # # # ~Reference

          # # # post.mean l-95% CI u-95% CI eff.samp
# # # Reference      2077    213.4     5054    25400

               # # # ~species

        # # # post.mean l-95% CI u-95% CI eff.samp
# # # species     887.4    120.7     2263    25400

 # # # R-structure:  ~units

      # # # post.mean l-95% CI u-95% CI eff.samp
# # # units      1339     1027     1683    25400

 # # # Location effects: devbysurv ~ bmrbysurv:cyclenr_char + cyclenr_char 

                        # # # post.mean l-95% CI u-95% CI eff.samp  pMCMC  
# # # (Intercept)               -13.367  -55.701   26.617    25400 0.5056  
# # # cyclenr_char1              15.161    2.285   28.432    25400 0.0213 *
# # # bmrbysurv:cyclenr_char0  -104.965 -187.625  -23.421    25400 0.0111 *
# # # bmrbysurv:cyclenr_char1   -50.822 -129.462   26.764    25400 0.1882  
# # # ---
# # # Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

# # # control for autocorrelation

autocorr(model1_no$VCV)
```
### Run the best model with phylogeny using mulTree

set the directory, where you wish to store the outputs
```{r, eval = F}
setwd("C:/Users/meeT55/Desktop/Julia/analysis_with_bmr/LOG-transformed/after_Martin/bmr_no_shorebirds")

# load phylogenetic trees without shorebirds

load("C:\\Users\\meeT55\\Desktop\\Julia\\data\\all_birds_tree_subset.RData")
tr_full<-all_birds_tree_subset

# set the model structure and parameters

mulTree_data <- as.mulTree(data = Circannual_cycles_no_shorebirds, tree = tr_full,    taxa = "animal",  rand.term=~uniqueid+Trait+Reference+animal+species)
formula_fintest<-devbysurv~bmrbysurv:cyclenr_char+cyclenr_char
mulTree.parameters<-c(5000000, 5000, 100000)

# run mulTree

bmr_n_shore<-mulTree(mulTree.data = mulTree_data, formula = formula_fintest, priors = pri5_1, pr=TRUE, parameters = mulTree.parameters, output = "bmr_n_shore", chains=8, parallel='SOCK')
```

## EXTRACT mulTree OUTPUT 
### we use Model 1 with shorebirds as an example

it is very likely that you will have to do it in a new R window. Therefore, again:
### load the packages
```{r, eval = F}
library(snow)
library(mulTree)
```
### set work directory, where the model output is stored
```{r, eval = F}
setwd("C:/Users/meeT55/Desktop/Julia/analysis_with_bmr/LOG-transformed/after_Martin/surv_with_shorebirds")
```

### Look if all the chains have converged

this function gets convergence parameters from 'conv' files, automatically saved in the same directory as model outputs
```{r, eval = F}
extract_conv<-function(model_name) {
  Files<-list.files( pattern=paste0(model_name,'.*_conv'))
  cat(length(Files), 'files found\n')
  Res<-c()
  for(i in 1:length(Files)){
  mod<-load(Files[i])
  treename<-Files[i]
  res<-unlist(converge.test)<1.1
Res1<-c(treename,res)
Res<-rbind(Res,Res1)
  }
  return(Res)
  }
  ```
### control for convergence
```{r, eval = F}
conv_mother<-extract_conv('surv_w_shore')
```
## Collect mulTree outputs

you can first look at one of the models output, to see its structure
```{r, eval = F}
load("C:/Users/meeT55/Desktop/Julia/analysis_with_bmr/LOG-transformed/after_Martin/bmr_no_shorebirds/bmr_n_shore-tree1_chain1.rda")
summary(model)
```
we need p-values (pMCMC) for the fixed parameters: 
pMCMC_cycle: intercept between the transitory (baseline) and full cycle (cyclenr_char1)
pMCMC_surv_cycle_0: slope for the transitory cycle
pMCMC_surv_cycle_1: slope for the full cycle

### function extracting pMCMC
```{r, eval = F}
extract_pMCMC<-function(model_name) {
  Files<-list.files( pattern=paste0(model_name,'.*_chain'))
  cat(length(Files), 'files found\n')
	pMCMC_cycle<-c()
	pMCMC_slope_cycle_0<-c()
	pMCMC_slope_cycle_1<-c()

     for(i in 1:length(Files)){
  mod<-load(Files[i])
  pMCMC_cycle <-c(pMCMC_cycle, summary(model)$solutions[18])
  pMCMC_slope_cycle_0 <-c(pMCMC_slope_cycle_0, summary(model)$solutions[19])
  pMCMC_slope_cycle_1 <-c(pMCMC_slope_cycle_1, summary(model)$solutions[20])

  }
  Res<-list(pMCMC_cycle=pMCMC_cycle,pMCMC_slope_cycle_0=pMCMC_slope_cycle_0,pMCMC_slope_cycle_1=pMCMC_slope_cycle_1)
  return(Res)
}
```
### extract pMCMC for the model
```{r, eval = F}
pMCMC<-extract_pMCMC('bmr_n_shore')
summary(pMCMC$pMCMC_cycle)
summary(pMCMC$pMCMC_slope_cycle_0)
summary(pMCMC$pMCMC_slope_cycle_1)
```
### function extracting fixed effects
```{r, eval = F}
extract_fixed<-function(model_name) {
  Files<-list.files( pattern=paste0(model_name,'.*_chain'))
  cat(length(Files), 'files found\n')
	Intercept<-c()
	cycle_nr_1<-c()
	slope_cycle_0<-c()
	slope_cycle_1<-c()
  
     for(i in 1:length(Files)){
  mod<-load(Files[i])
  Intercept <-c(Intercept, model$Sol[,1])
  cycle_nr_1 <-c(cycle_nr_1, model$Sol[,2])
  slope_cycle_0 <-c(slope_cycle_0, model$Sol[,3])
  slope_cycle_1 <-c(slope_cycle_1, model$Sol[,4])
  }
  Res<-list(Intercept=Intercept, cycle_nr_1=cycle_nr_1,slope_cycle_0=slope_cycle_0,slope_cycle_1=slope_cycle_1)
  return(Res)
}
```
### extract fixed effects
```{r, eval = F}
fixed_effects<-extract_fixed('surv_w_shore')
summary(fixed_effects$Intercept)
summary(fixed_effects$cycle_nr_1)
summary(fixed_effects$slope_cycle_0)
summary(fixed_effects$slope_cycle_1)

quantile(fixed_effects$Intercept,c(0.025,0.975))
quantile(fixed_effects$cycle_nr_1,c(0.025,0.975))
quantile(fixed_effects$slope_cycle_0,c(0.025,0.975))
quantile(fixed_effects$slope_cycle_1,c(0.025,0.975))
```

### function extracting random effects
```{r, eval = F}
extract_random<-function(model_name) {
  Files<-list.files( pattern=paste0(model_name,'.*_chain'))
  cat(length(Files), 'files found\n')
  uniqueid_post<-c()
  Trait_post<-c()
  Reference_post<-c()
  animal_post<-c()
  species_post<-c()
  units_post<-c()
     for(i in 1:length(Files)){
  mod<-load(Files[i])
  uniqueid_post <-c(uniqueid_post, model$VCV[,1])
  Trait_post <-c(Trait_post, model$VCV[,2])
  Reference_post <-c(Reference_post, model$VCV[,3])
  animal_post <-c(animal_post, model$VCV[,4])
  species_post <-c(species_post, model$VCV[,5])
  units_post <-c(units_post, model$VCV[,6])
   
  }
  Res<-list(uniqueid_post=uniqueid_post, Trait_post=Trait_post,Reference_post=Reference_post,animal_post=animal_post, species_post=species_post,units_post=units_post)
  return(Res)
}
```
### extract random effects
```{r, eval = F}
random_effects<-extract_random('bmr_n_shore')
summary(random_effects$uniqueid_post)
summary(random_effects$Trait_post)
summary(random_effects$Reference_post)
summary(random_effects$animal_post)
summary(random_effects$species_post)
summary(random_effects$units_post)

quantile(random_effects$uniqueid_post,c(0.025,0.975))
quantile(random_effects$Trait_post,c(0.025,0.975))
quantile(random_effects$Reference_post,c(0.025,0.975))
quantile(random_effects$animal_post,c(0.025,0.975))
quantile(random_effects$species_post,c(0.025,0.975))
quantile(random_effects$units_post,c(0.025,0.975))
```

### calculate phylogenetic signal
```{r, eval = F}
phyl_sig<-var(random_effects$animal_post)/(var(random_effects$uniqueid_post)+var(random_effects$Trait_post)+var(random_effects$Reference_post)+var(random_effects$animal_post)+var(random_effects$species_post)+var(random_effects$units_post))

lambda<-var(random_effects$animal_post)/(var(random_effects$animal_post)+var(random_effects$species_post))
```

