---
title: "Appendix S6. Testing for the correlation of free-running circannual period lengths with survival using MCMCglmm and mulTree"
author: "Julia Karagicheva, Eldar Rakhimberdiev, Anatoly Saveliev & Theunis Piersma"
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhf{}
- \fancyhead[LO,LE]{Functional Ecology}
- \fancyfoot[LO,LE]{Annual chronotypes functionally link life histories and life cycles in birds, Karagicheva et al.}
- \fancyfoot[RE,RO]{\thepage}
---

Supplementary online material S6 to Julia Karagicheva, Eldar Rakhimberdiev, Anatoly Saveliev & Theunis Piersma XXXX _Annual chronotypes functionally link life histories and life cycles in birds_. - Journal of Functional Ecology 000: 000-000.

The R code for the presented in the paper analyses, assessing the relationships between deviations of circannual periods from 365 days (circannual deviations) and species-specific annual survival rates. 

Install required packages

```{r, eval = F}
install.packages('snow')
if(!require(devtools)) install.packages("devtools")
library(devtools)
install_github("TGuillerme/mulTree", ref = "release")
```

you might need to install some packages manually
```{r, eval = F}
install.packages(c("MCMCglmm", "coda", "hdrcde", "snow", "ape", "corpcor", "curl"))
```

Load required packages
```{r, eval = F}
library(snow)
library(mulTree)
```

### PREPARE THE DATA 
Download data from GitHub

load table with circannual cycles and deviations of circannual period from 365 days (main table "Circannual_cycles_birds_data")

```{r, eval = F}
Circannual_cycles_birds_data<-read.csv('https://git.io/v5GIF',stringsAsFactors=F)
```

create unique ID 
```{r, eval = F}
Circannual_cycles_birds_data$uniqueid<-paste(Circannual_cycles_birds_data$species,
     Circannual_cycles_birds_data$Reference,
     Circannual_cycles_birds_data$BirdID_char,sep="_")
```

make cycle number character
```{r, eval = F}
Circannual_cycles_birds_data$cyclenr_char<-
     as.character(Circannual_cycles_birds_data$cyclenr_char)
```

load table "Survival_birds_data" with species-specific survival and calculate median per species
```{r, eval = F}
Survival_birds_data<-read.csv('https://git.io/v5Gtq',stringsAsFactors=F)
surv_sp<-aggregate(Survival_birds_data$survival,
      by=list(Survival_birds_data$species),median)
```

add survival values to the main table
```{r, eval = F}
Circannual_cycles_birds_data$surv<-NA
for(i in 1:nrow(Circannual_cycles_birds_data)){
     Circannual_cycles_birds_data$surv[i]<-
        unique(surv_sp[surv_sp$Group.1 %in%
        Circannual_cycles_birds_data$species[i],]$x)
}
```

load table "BMR_birds_data.csv" with species-specific BMR  and calculate median per species
```{r, eval = F}
BMR_birds_data<-read.csv('https://git.io/v5ZSb',stringsAsFactors=F)
bmr_sp_median<-aggregate(BMR_birds_data$BMR_W,
                 by=list(BMR_birds_data$species),median)
```

add BMR values to the main table
```{r, eval = F}
Circannual_cycles_birds_data$bmr<-NA
for(i in 1:nrow(Circannual_cycles_birds_data)){
    Circannual_cycles_birds_data$bmr[i]<-
       unique(bmr_sp_median[bmr_sp_median$Group.1==
       Circannual_cycles_birds_data$species[i],]$x)
}
```

Calculate residuals of survival on BMR
```{r, eval = F}
Circannual_cycles_birds_data$resbmr<-
residuals(lm(log(surv)~log(bmr),data=Circannual_cycles_birds_data))
```

Calculate residuals of circannual deviation on BMR
```{r, eval = F}
Circannual_cycles_birds_data$res_dev_bmr<-
     residuals(lm(deviation~log(bmr),data=Circannual_cycles_birds_data))
```

### RUN THE MODELS
#### set priors
```{r, eval = F}
p.var<-var(Circannual_cycles_birds_data$res_dev_bmr,na.rm=TRUE)
pri4_1 <- list(R = list(V = matrix(p.var/2), nu = 1), 
     G = list(G1 = list(V = matrix(p.var/2), nu = 1),
     G2 = list(V = matrix(p.var/2), nu = 1), 
     G3=list(V = matrix(p.var/2), nu = 1),
     G4=list(V = matrix(p.var/2), nu = 1)))
pri5_1 <- list(R = list(V = matrix(p.var/2), nu = 1), 
     G = list(G1 = list(V = matrix(p.var/2), nu = 1),
     G2 = list(V = matrix(p.var/2), nu = 1), 
     G3=list(V = matrix(p.var/2), nu = 1),
     G4=list(V = matrix(p.var/2), nu = 1),
     G5=list(V = matrix(p.var/2), nu = 1)))
```

#### run MCMCglmm without phylogeny (see Table S7 in Appendix for the best selected model output)
```{r, eval = F}
model1 <- MCMCglmm(res_dev_bmr~resbmr:cyclenr_char+cyclenr_char,
                   random= ~uniqueid+Trait+Reference+species,
		   data = Circannual_cycles_birds_data, prior=pri4_1,
		   nitt=130000*10,thin=10*5,burnin=3000*10,pr=TRUE,verbose=F,nodes="ALL")

model2 <- MCMCglmm(res_dev_bmr~resbmr+cyclenr_char,
                   random= ~uniqueid+Trait+Reference+species,
		   data = Circannual_cycles_birds_data, prior=pri4_1,
		   nitt=130000*10,thin=10*5,burnin=3000*10,pr=TRUE,verbose=F,nodes="ALL")

model1$DIC #the best model
#4000.682 
model2$DIC
#4008.936
```

#### get phylogenetic data
Download phylogenetic trees (containing shorebirds), directly from http://birdtree.org/ or just download the 100 trees used in the presented analysis from https://git.io/v5ZxG:
```{r, eval = F}
download.file("https://git.io/v5ZxG", "with_shorebirds_tree_subset.RData",mode="wb")
# load the phylogenetic trees subset
load("with_shorebirds_tree_subset.RData")
tr_full<-with_shorebirds_tree_subset

# select one phylogenetic tree for model selection in MCMCglmm
mulTree_data <- as.mulTree(data = Circannual_cycles_birds_data,
                           tree = tr_full, taxa = "animal",
			   rand.term=~uniqueid+Trait+Reference+species+animal)

ped<-mulTree_data$phy[[1]]
```

#### run MCMCglmm with phylogeny (see Table S7 in Appendix for the best selected model output)
```{r, eval = F}
model1_pedigree <- MCMCglmm(res_dev_bmr~resbmr:cyclenr_char+cyclenr_char,
                            random= ~uniqueid+Trait+Reference+species+animal,
			    data= Circannual_cycles_birds_data, pedigree=ped,prior=pri5_1,
			    nitt=130000*10,thin=10*5,burnin=3000*10,pr=TRUE,verbose=F,nodes="ALL")

model2_pedigree <- MCMCglmm(res_dev_bmr~resbmr+cyclenr_char,
                            random= ~uniqueid+Trait+Reference+species+animal,
			    data= Circannual_cycles_birds_data, pedigree=ped,prior=pri5_1,
			    nitt=130000*10,thin=10*5,burnin=3000*10,pr=TRUE,verbose=F,nodes="ALL")

model1_pedigree$DIC
# 4000.948
model2_pedigree$DIC
# 4009.138
```

The best model
```{r, eval = F}
summary(model1_pedigree)
```

check autocorrelation
```{r, eval = F}
autocorr(model1_pedigree$VCV)
```

#### run the best model with phylogeny using mulTree
set the directory, where you wish to store the outputs
##### set the model structure and parameters
```{r, eval = F}
mulTree_data <- 
as.mulTree(data = Circannual_cycles_birds_data, tree = tr_full,
     taxa = "animal", rand.term=~uniqueid+Trait+Reference+animal+species)
formula_fintest<-res_dev_bmr~resbmr:cyclenr_char+cyclenr_char
mulTree.parameters<-c(5000000, 5000, 100000)
```

##### run mulTree
```{r, eval = F}
surv_w_shore<-mulTree(mulTree.data = mulTree_data, 
     formula = formula_fintest, priors = pri5_1, pr=TRUE, 
     parameters = mulTree.parameters, output = "surv_w_shore",
     chains=8, parallel='SOCK')
```

### EXTRACT OUTPUTS 
it is very likely that you will have to do it in a new R window. Therefore, again:
```{r, eval = F}
# load the packages
library(snow)
library(mulTree)
#... and set the work directory, where the model output is stored
```

#### check whether all the chains have converged
this function gets convergence parameters from 'conv' files, automatically saved in the same directory as model outputs
```{r, eval = F}
extract_conv<-function(model_name) {
  Files<-list.files( pattern=paste0(model_name,'.*_conv'))
  cat(length(Files), 'files found\n')
  Res<-c()
  for(i in 1:length(Files)){
     mod<-load(Files[i])
     treename<-Files[i]
     res<-unlist(converge.test)<1.1
     Res1<-c(treename,res)
     Res<-rbind(Res,Res1)
  }
  return(Res)
}
```

#### control for convergence
```{r, eval = F}
conv_mother<-extract_conv('surv_w_shore')
```

#### collect mulTree outputs
you can first load one of the models' output, to see its structure

we need p-values (pMCMC) for the fixed parameters: 
pMCMC_cycle: intercept between the transitory (baseline) and full cycle (cyclenr_char1)
pMCMC_surv_cycle_0: slope for the transitory cycle
pMCMC_surv_cycle_1: slope for the full cycle

#### function extracting pMCMC
```{r, eval = F}
extract_pMCMC<-function(model_name) {
  Files<-list.files( pattern=paste0(model_name,'.*_chain'))
  cat(length(Files), 'files found\n')
	pMCMC_cycle<-c()
	pMCMC_slope_cycle_0<-c()
	pMCMC_slope_cycle_1<-c()

  for(i in 1:length(Files)){
     mod<-load(Files[i])
     pMCMC_cycle <-c(pMCMC_cycle, summary(model)$solutions[18])
     pMCMC_slope_cycle_0 <-c(pMCMC_slope_cycle_0, summary(model)$solutions[19])
     pMCMC_slope_cycle_1 <-c(pMCMC_slope_cycle_1, summary(model)$solutions[20])
  }
  Res<-list(pMCMC_cycle=pMCMC_cycle,pMCMC_slope_cycle_0=pMCMC_slope_cycle_0,
       pMCMC_slope_cycle_1=pMCMC_slope_cycle_1)
  return(Res)
}
```

#### extract pMCMC for the model
```{r, eval = F}
pMCMC<-extract_pMCMC('surv_w_shore')
summary(pMCMC$pMCMC_cycle)
summary(pMCMC$pMCMC_slope_cycle_0)
summary(pMCMC$pMCMC_slope_cycle_1)
```

#### function extracting fixed effects
```{r, eval = F}
extract_fixed<-function(model_name) {
  Files<-list.files( pattern=paste0(model_name,'.*_chain'))
  cat(length(Files), 'files found\n')
	Intercept<-c()
	cycle_nr_1<-c()
	slope_cycle_0<-c()
	slope_cycle_1<-c()
 
  for(i in 1:length(Files)){
     mod<-load(Files[i])
     Intercept <-c(Intercept, model$Sol[,1])
     cycle_nr_1 <-c(cycle_nr_1, model$Sol[,2])
     slope_cycle_0 <-c(slope_cycle_0, model$Sol[,3])
     slope_cycle_1 <-c(slope_cycle_1, model$Sol[,4])
  }
  Res<-list(Intercept=Intercept, cycle_nr_1=cycle_nr_1,
       slope_cycle_0=slope_cycle_0,slope_cycle_1=slope_cycle_1)
  return(Res)
}
```

#### extract fixed effects
```{r, eval = F}
fixed_effects<-extract_fixed('surv_w_shore')
summary(fixed_effects$Intercept)
summary(fixed_effects$cycle_nr_1)
summary(fixed_effects$slope_cycle_0)
summary(fixed_effects$slope_cycle_1)

quantile(fixed_effects$Intercept,c(0.025,0.975))
quantile(fixed_effects$cycle_nr_1,c(0.025,0.975))
quantile(fixed_effects$slope_cycle_0,c(0.025,0.975))
quantile(fixed_effects$slope_cycle_1,c(0.025,0.975))
```

#### function extracting random effects
```{r, eval = F}
extract_random<-function(model_name) {
  Files<-list.files( pattern=paste0(model_name,'.*_chain'))
  cat(length(Files), 'files found\n')
  uniqueid_post<-c()
  Trait_post<-c()
  Reference_post<-c()
  animal_post<-c()
  species_post<-c()
  units_post<-c()
  for(i in 1:length(Files)){
     mod<-load(Files[i])
     uniqueid_post <-c(uniqueid_post, model$VCV[,1])
     Trait_post <-c(Trait_post, model$VCV[,2])
     Reference_post <-c(Reference_post, model$VCV[,3])
     animal_post <-c(animal_post, model$VCV[,4])
     species_post <-c(species_post, model$VCV[,5])
     units_post <-c(units_post, model$VCV[,6])
  }
  Res<-list(uniqueid_post=uniqueid_post, Trait_post=Trait_post,
       Reference_post=Reference_post,animal_post=animal_post, 
       species_post=species_post,units_post=units_post)
  return(Res)
}
```

#### extract random effects
```{r, eval = F}
random_effects<-extract_random('surv_w_shore')
summary(random_effects$uniqueid_post)
summary(random_effects$Trait_post)
summary(random_effects$Reference_post)
summary(random_effects$animal_post)
summary(random_effects$species_post)
summary(random_effects$units_post)

quantile(random_effects$uniqueid_post,c(0.025,0.975))
quantile(random_effects$Trait_post,c(0.025,0.975))
quantile(random_effects$Reference_post,c(0.025,0.975))
quantile(random_effects$animal_post,c(0.025,0.975))
quantile(random_effects$species_post,c(0.025,0.975))
quantile(random_effects$units_post,c(0.025,0.975))
```

### CALCULATE PHYLOGENETIC SIGNAL AND REPEATABILITIES
```{r, eval = F}
# phylogenetic signal and intra-specific repeatability
phyl_sig<-var(random_effects$animal_post)/
          (var(random_effects$uniqueid_post) + var(random_effects$Trait_post) + 
	   var(random_effects$Reference_post) + var(random_effects$animal_post) + 
	   var(random_effects$species_post)+var(random_effects$units_post))
# phyl_sig = 0.423506

lambda<-var(random_effects$animal_post)/
        (var(random_effects$animal_post)+var(random_effects$species_post))
#lambda = 0.812362

#Intra-specific repeatability
repeatab<-var(random_effects$species_post)/
          (var(random_effects$uniqueid_post)+var(random_effects$Trait_post) + 
	   var(random_effects$Reference_post)+ var(random_effects$animal_post) +
	   var(random_effects$species_post)+var(random_effects$units_post))
# repeatab = 0.09782072

repeatab_adjusted<-var(random_effects$species_post)/
                    (var(random_effects$species_post)+var(random_effects$units_post))
# repeatab_adjusted = 0.9517922
```
